Reviewer comments from PeerJ

### 2015-02-19 PeerJ_trackchanges.tex
Edits
figure numbers

\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations. DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb). Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig. S1 and \track{Fig. 1}, See Supplemental Materials with the online version of this article). Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Table S2, See Supplemental Materials with the online version of this article). Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

@@

\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened. Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig1Map}). Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes. We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}). Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).

@@

\textbf{Evidence of introgression---}The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize. Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte. To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome 1 and specifically in the \emph{tb1} region. If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element. We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion. In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).

The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) \citep{Pyhajarvi2013} and the maize 282 diversity panel \citep{Cook2012}. 
The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants which would arise as the result of recent admixture between populations. If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the ``maize" cluster in our STRUCTURE runs, not the ``teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig. \ref{Fig3Structure}).

\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping. 
We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion. 
We also measured culm diameter, a phenotype that differs between maize and teosinte but has not been shown to be affected by the \emph{Hopscotch} insertion \citep{Briggs2007}. Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data. If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion, the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype. 
Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
Our plantings produced 82 homozygotes for the \emph{Hopscotch} insertion at \emph{tb1}, 104 heterozygotes, and 20 homozygotes lacking the insertion; these numbers do not deviate from expectations of Hardy-Weinberg equilibrium. 
After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig. \ref{Fig4Boxplots}), tiller number, or culm diameter. 
Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.

@@

While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results do not suggest introgression from maize at the \emph{tb1} locus, and are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into \emph{mexicana} around domestication loci. 
Clustering in our NJ trees does not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.  Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig. \ref{Fig3Structure}, Table \ref{Table4Q}). Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in the Jalisco cluster of our \emph{parviglumis} populations.

@@

%Figure 2 Map 
%--------------------------------------------
\begin{figure*}[!t]
\begin{center}
   \includegraphics[width=0.8\textwidth]{Fig2MapBalsas.png}
   \end{center}
    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals. Size of circles reflects number of individuals sampled. The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.} 
\label{Fig1Map}
\end{figure*}
%--------------------------------------------
\clearpage

>>>>

\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations. DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb). Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig. S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article). Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Table S2, See Supplemental Materials with the online version of this article). Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

@@

\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened. Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig2Map}). Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes. We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}). Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).

@@

\textbf{Evidence of introgression---}The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize. Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte. To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome 1 and specifically in the \emph{tb1} region. If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element. We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.\ref{Fig3LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion. In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).

The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) \citep{Pyhajarvi2013} and the maize 282 diversity panel \citep{Cook2012}. 
The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants which would arise as the result of recent admixture between populations. If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the ``maize" cluster in our STRUCTURE runs, not the ``teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig. \ref{Fig4Structure}).

\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping. 
We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion. 
We also measured culm diameter, a phenotype that differs between maize and teosinte but has not been shown to be affected by the \emph{Hopscotch} insertion \citep{Briggs2007}. Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data. If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion, the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype. 
Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
Our plantings produced 82 homozygotes for the \emph{Hopscotch} insertion at \emph{tb1}, 104 heterozygotes, and 20 homozygotes lacking the insertion; these numbers do not deviate from expectations of Hardy-Weinberg equilibrium. 
After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig. \ref{Fig5Boxplots}), tiller number, or culm diameter. 
Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.

@@

While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results do not suggest introgression from maize at the \emph{tb1} locus, and are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into \emph{mexicana} around domestication loci. 
Clustering in our NJ trees does not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.  Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig. \ref{Fig4Structure}, Table \ref{Table4Q}). Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in the Jalisco cluster of our \emph{parviglumis} populations.

@@

%Figure 2 Map 
%--------------------------------------------
\begin{figure*}[!t]
\begin{center}
   \includegraphics[width=0.8\textwidth]{Fig2MapBalsas.png}
   \end{center}
    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals. Size of circles reflects number of individuals sampled. The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.} 
\label{Fig2Map}
\end{figure*}
%--------------------------------------------
\clearpage
